---
title: "lmap, modify, predicate functionals"
author: "Jack Penzer"
date: "2023-07-16"
output: 
  rmdformats::readthedown:
    code_folding: show
    toc_depth: 2
    use_bookdown: true
# template:
#   bootstrap: 5
#   bootswatch: united
#   theme: arrow-light
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```



```{r}
library(purrr)
library(dplyr)
```


Last week Sarah took us through `map2()` and `pmap()`, this weekwe'll look at `lmap()`, `modify()` and some predicate functionals 

Reminder: predicate functionals return TRUE or FALSE

# lmap

What is it?

```{r}
?lmap
```

"Apply a function to list-elements of a list"

What on earth is a list-element of a list? An element inside a list that is a list itself

```{r}
set.seed(1014)
maybe_rep <- function(x) {
  n <- rpois(1, 2)
  set_names(rep_len(x, n), paste0(names(x), seq_len(n)))
}



# The output size varies each time we map f()
(x <- list(a = 1:4, b = letters[5:7], c = 8:9, d = letters[10]))
```


lmap(), lmap_at() and lmap_if() are similar to map(), map_at() and map_if(), except instead of mapping over .x[[i]], they instead map over .x[i]. Recall that [[]] returns object itself, [] returns a list with object in. lmap only takes a list as input (not vectors)

```{r}
lmap(list(1:5), maybe_rep) # Fine
```

Doesn't take vectors:

```{r, eval = FALSE}
lmap(c(1:5), maybe_rep) # Error
```

From the docs:
The lmap function has several advantages:

- It makes it possible to work with functions that exclusively take a list.
- It allows .f to access the attributes of the encapsulating list, like names().
- It allows .f to return a larger or small list than it receives changing the size of the output.

## How does it work?

```{r}
set.seed(187)
x |> lmap(maybe_rep)
```

```{r}
set.seed(123)
x |> map(maybe_rep)
```

`lmap()` calls `lmap_helper()`
```{r}
lmap_helper <- function (.x, .ind, .f, ..., .else = NULL, .purrr_error_call = caller_env()) 
{
    .f <- rlang::as_function(.f, call = .purrr_error_call)
    if (!is.null(.else)) {
        .else <- rlang::as_function(.else, call = .purrr_error_call)
    }
    out <- vector("list", length(.x))
    for (i in seq_along(.x)) {
        if (.ind[[i]]) {
            res <- .f(.x[i], ...)
        }
        else if (is.null(.else)) {
            res <- .x[i]
        }
        else {
            res <- .else(.x[i], ...)
        }
        if (!is.list(res)) {
            cli::cli_abort("{.code .f(.x[[{i}]])} must return a list, not {.obj_type_friendly {res}}.", 
                call = .purrr_error_call)
        }
        out[[i]] <- res
    }
    if (is.data.frame(.x)) {
        out <- lapply(out, as.data.frame)
        list_cbind(out)
    }
    else {
        list_flatten(out)
    }
}
```

## Notes on documentation

`...` Additional arguments passed on to the mapped function.

 We now generally recommend against using ... to pass additional (constant) arguments to .f. Instead use a shorthand anonymous function:

<!-- # Instead of -->
<!-- x |> map(f, 1, 2, collapse = ",") -->
<!-- # do: -->
<!-- x |> map(\(x) f(x, 1, 2, collapse = ",")) -->


## Variants

### lmap_if

Similar to lmap, except it has two new arguments .p, which is a predicate function (a function that returns TRUE/FALSE) and .else which is what to do when the predicate function returns FALSE.

```{r}
# Or only where a condition is satisfied
x |> lmap_if(is.character, maybe_rep) |> str()
```

### lamp_at

Similar to `lmap_if()`, but you refer to the elements via a vector rather than a predicate function.

```{r}
# We can apply f() on a selected subset of x
x |> lmap_at(c("a", "d"), maybe_rep) |> str()
```

# modify

Key difference between the `modify()` and `map()` family is that `map` and its variants (chr, int, lgl etc.) always return the same data type. `modify` returns the same data type as the input.

`Modify` takes a vector, list, or data frame object as input.

```{r}
modify(c(1:5), ~ .x + 1)
modify(list(1:5), ~ .x +2)
```

When we feed modify a data frame as input, it returns a data frame as output

```{r}
iris %>%
  modify(as.character) %>%
  tibble::tibble()
```

Can replace some of the mutate(across()) stuff people often want to do :D

## modify_if

```{r}
iris %>%
  modify_if(.p = is.factor, .f = as.character) %>%
  tibble::tibble()
```

versus

```{r}
iris %>%
  map_if(is.factor, as.character) 
```


## modify_at

Similar to lmap_at etc. we can use a vector of indices or names

```{r}
iris %>%
  modify_at(c(1, 2, 3, 4), as.integer) %>%
  tibble::tibble()
```

modify_depth
```{r}
x <- list(a = list(foo = 1:2, bar = 3:4), b = list(baz = 5:6))
str(x)

modify_depth(.x = x, .depth = 2, .f = function(x) x + 1)
```



## modify_tree

```{r}
x <- list(list(a = 2:1, c = list(b1 = 2), b = list(c2 = 3, c1 = 4)))
x |> str()


# Transform each leaf
x |> modify_tree(leaf = \(x) x + 100) |>  str()


# Recursively sort the nodes
sort_named <- function(x) {
  nms <- names(x)
  if (!is.null(nms)) {
    x[order(nms)]
  } else {
    x
   }
}
x |> modify_tree(post = sort_named) |> str()
```


# predicate functionals

Predicate functionals are functions which return TRUE or FALSE.

## detect

Detect accepts a list or vector and will find the value of the first match. 

Similar to map it takes .x and .f as inputs, as well as `...`, docs say '	
Additional arguments passed on to .p.' but I'm assuming it means .f, or they meant to use .p as the function rather than .f.

```{r}
vec <- c(5:20)
detect(vec, ~ .x > 10) 
detect_index(vec, ~.x > 10)
```

We get a new (I think?) argument, .dir which asks which side of the vector and takes 'forward' or 'backward as inputs. 

```{r}
detect(vec, ~ .x > 10, .dir = "backward")
```
Returns 20 as we'd expect, as from the end of the vector, the first element which is > 10 is the first element - 20.

```{r}
detect_index(vec, ~ .x > 10, .dir = "backward")
```

returns 16 as we'd expect, as len(5:20) == 16

## every, some, none

They work just like `any()` and `all()`, with the negated version of all being none.

```{r}
x <- list(0:10, 5.5)
x |> every(is.numeric)
#> [1] TRUE
```


```{r}
x |> every(is.integer)
#> [1] FALSE
```


```{r}
x |> some(is.integer)
#> [1] TRUE
```


```{r}
x |> none(is.character)
#> [1] TRUE
```

## has_element

Function simply checks if a list or atomic vector contains an object. 

```{r}
x <- list(1:10, 5, 9.9)
has_element(x, 1:10)
```

Essentially this is checking if there is `some()` element which is identical to any of the other elements:
```{r eval = FALSE}
function (.x, .y) 
{
    some(.x, identical, .y)
}
```

Ran out of time in prep !: Next week continue at head_while and tail_while
```{r}
list(x = sample(c(TRUE, FALSE), 100, replace = TRUE), y = 1:100) |>
  list_transpose(simplify = FALSE) |>
  modify_if("x", \(l) list(x = l$x, y = l$y * 100)) |>
  list_transpose()
```

