---
title: "Week 2, Jenny Bryan's Tutorial"
author: "Jack Penzer"
date: "2023-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This week we'll be skimming over Jenny Bryan's excellent set of {purrr} tutorials. We'll be doing our best to condense them into an hour - as a primer for getting down and dirty with the docs next week.

```{r}
library(purrr)
library(repurrrsive)
```

# Background Basics

As we'll often be working with lists and vectors, it'll be helpful to have a quick refresher:

## Vectors & lists

Recall in R, the main difference between a vector and a list, is that a vector can only store one type of data structure whereas lists can store many.

```{r}
c(1, 2, 3)
c("1", 2 , 3)#2 and 3 are co-erced to characters
c(1, 2, "3")  #bigger memory allocation to string so coerces to string=
c(1,  2, "three") # 1 and 2 are co-erced to characters

list(1, 2, "three") #Nothing gets co-erced
list("1", "2", "three")
```

We can index vectors using [] notation or [[]]]

```{r}
my_vec <- c(3, 2, 1)
my_vec[1]
my_vec[2]
my_vec[3]

#We can drop elements with negatives:
my_vec[-2] #keeps 1st & 3rd elements
my_vec[-c(1, 2)] #just keeps the 3rd element
```

Both vectors and lists can be named:

```{r}
named_vec <- c(one = 1, two = 2, three = 3)

named_list <- list(one = 1, two = 2, three = 3)
```

We can index the names of vectors using [] syntax, but not \$

```{r}
named_vec["one"]
named_vec[c("one", "two", "three")]
named_vec$one # error
```

But we can use \$ for lists:

```{r}
class(named_list[1])
class(named_list[[1]])
```

When indexing lists, [] will return a list, [[]] will return the element itself (it's important to distinguish the list from its elements)

```{r}
named_list[["one"]]
named_list[[1]]
```

## str()

Before you can work efficiently with a list in the wild, you'll need to understand its structure. Jenny Bryan gives three things to be on the lookout for:

-   What length is the list, what depth do its components have?

-   Do the components have the same data structure, i.e. are they homogeneous?

-   Do the components have names?

It's important to keep note of the answers to all three.

Jenny Bryan's repurrrsive package will provide us with a few helpful lists (consider re-using these in future weeks when demonstrating concepts!)

the got_chars list has a bunch of nested lists for Game of Thrones characters. We can use str() to inspect the list

```{r}
got <- repurrrsive::got_chars
got
length(got)
str(got)
length(got[[1]])
```


## Lists sometimes have ugly properties

When dealing with vectors, many operations are vectorised (apply to each element in the vector) right out of the box:

```{r}
c(1, 2, 3) ^ 2
```

But this isn't true with lists:

```{r}
list(1, 2, 3) ^ 2
```

# List extraction -  Game of Thrones characters
 and we'll set list.len = 5 so as not to spam the console:

```{r}
str(got, list.len = 2)
```

We can also look at a specific got character:

```{r}
(first_character <- got[[1]])
```

Then index by names:

```{r}
names(first_character)
first_character$name #let R STudio autocorrect help you!
```

Let's say we want to get the names of the first 10 characters out of the got object, it would be nice if we could just index 1:10, and type \$name:

```{r}
got[[1:10]]$name
```

So that doesn't work, in fact we can't even get the 1:10 index this way:

```{r}
got[[1:10]]
```

Using one square bracket won't help, either:

```{r}
got[1:10]
got[1:10]$name
```

## Map to the rescue

It's also important to remember `map()` always returns a list. However, there are `map_` variants which return other types. When not comfortable with lists, this predictable fact about `map` might seem ugly or inconvenient, but over time it becomes quite the opposite.

How do we use map to get the names of the first 10 characters? Because each element in got[1:10] has a $name object associated with it, we can use this to our advantage:

```{r}
map(got[1:10], "name")
map(got[1:10], "born")
map(got[1:10], "tower")
map(got, "name")
map(got, `[`, "name")
```

What data structure is the output?

When using map for extraction purposes, we can also use index. as the name is contained in the third index, we could access it by passing in the number 3:

```{r}
map(got[1:10], 3)
map(got[1:10], 4)
```


Compare the output of the previous call with the output of map_chr

```{r}
names_chr <- map_chr(got[1:10], "name")
class(names_chr)
```

What did map_chr do that `map` didn't?
```{r}
names_1_10 <- map(got[1:10], "name")
class(unlist(names_1_10))
```

# Mapping functions (going beyond extraction)

As we're learning {purrr}, we'll focus on how to apply functions to lists with {purrr}'s map functions. It'll be important to keep this structure in mind:

`map(your_list, your_function)`

when using ellipsis
`map(your_list, your_function, ...)` # where ... is your additional argument(s) to your_function

```{r}
dataframe <- data.frame()
dummy_function <- function(data, ..., na.rm = TRUE)
map(my_list, dummy_function, data = dataframe, na.rm = TRUE)
```

When you type ?map, you'll notice that map also takes '...', this is a way to feed in additional arguments to `your_function`. However, this list is not iterated over. This can start to get quite confusing and wordy, so let's look at an example.

## Existing functions

Demo lists:
```{r}
aliases <- set_names(map(got_chars, "aliases"), map_chr(got_chars, "name"))
(aliases <- aliases[c("Theon Greyjoy", "Asha Greyjoy", "Brienne of Tarth")])

```

We can choose to feed in a function that has already been defined prior to our map call. We'll define a function, my_fun and show how we do that.


```{r}
my_fun <- function(x) {
  paste(x, collapse = "|")
} 
paste(aliases$`Theon Greyjoy`, "|")
paste(aliases$`Asha Greyjoy`, "|")

map(aliases, my_fun)

map(aliases, my_fun, collapse = ";")
```

For each of the characters in our aliases list, we get a pasted version of their aliases, separated by |. If we wanted to separate by ";" we'd have to edit my_fun, as there is currently no way to tell my_fun what to separate by. Or, we could use an anonymous function and `...`  to feed additional arguments to paste.
```{r}
map(aliases, paste, collapse = ";")
```

Anonymous functions
```{r}
map(aliases, function(x) paste(x, collapse = ";"))
map(aliases, function(x, collapse) paste(x, collapse = "!"))
```

We could also get used to the anonymous function with a formula:
```{r}
map(.x = aliases, .f = ~ paste(.x, collapse = ";"))
```

We'll need the formula notation later when we get into `map2` and `pmap` territory. Plus it lets us build pipes etc. on the fly:
```{r}
map(aliases, ~ paste(.x, collapse = ";") %>%
      stringr::str_detect("Kinslayer")
    )
```


Exercises:
1. Create a list allegiances that holds the characters’ house affiliations.
```{r}
allegiances <- map(got, "allegiances")
```

2. Create a character vector nms that holds the characters’ names.
```{r}
nms <- map_chr(got, "name")
```

3. Apply the names in nms to the allegiances list via set_names.
```{r}
allegiances <- allegiances %>%
  setNames(nms)
```

4. Re-use the code from above to collapse each character’s vector of allegiances down to a string.
```{r}
map(allegiances, my_fun)
```

We said that any elements passed via ... would be used “as is”. Specifically they are not used in a vectorized fashion. What happens if you pass collapse = c(" | ", " * ")? Why is that?

1. Create a list allegiances that holds the characters’ house affiliations.
```{r}
names <- map(got_chars, "name")
names(got_chars) <- names

map(got_chars, "allegiances")
```

2.Create a character vector nms that holds the characters’ names.
```{r}
nms <- map(got_chars, "name")
```

3. Re-use the code from above to collapse each character’s vector of allegiances down to a string.
```{r}
allegiances <- map(got_chars, "allegiances")
allegieances <- allegiances %>%
  setNames(nms)
```

4. Re-use the code from above to collapse each character’s vector of allegiances down to a string.
```{r}
map(allegiances, paste, collapse = "|")
```

5. We said that any elements passed via ... would be used “as is”. Specifically they are not used in a vectorized fashion. What happens if you pass collapse = c(" | ", " * ")? Why is that?
```{r}
map(allegiances, paste, collapse = c("|", ", ", "* ")) # just takes the first element of the character vector?

map(allegiances, paste, collapse = c(",", "|", "* "))
```


## Recap

shortcut extraction with
map(list, "name")
or by index
map(list, index)

Existing functions, or anonymous functions, or bare function names, or anonymous function with formula notation.
map(aliases, my_fun)
map(aliases, function(x) paste(x, collapse = "|")) 
map(aliases, paste, collapse = "|")
map(aliases, ~ paste(.x, collapse = " | "))

## Multiple lists quick demo - map2 and pmap
```{r}
nms <- got_chars %>% 
  map_chr("name")
birth <- got_chars %>% 
  map_chr("born")

```


```{r}
my_fun <- function(x, y) paste(x, "was born", y)

map2_chr(.x = nms, .y = birth, .f = my_fun) %>% head()

map2_chr(.x = nms[16:18], .y = birth[16:18], ~ paste(.x, "was born", .y)) %>% tail()
```

```{r}
df <- got_chars %>% {
  tibble::tibble(
    name = map_chr(., "name"),
    aliases = map(., "aliases"),
    allegiances = map(., "allegiances")
  )
}


my_fun <- function(name, aliases, allegiances) {
  paste(name, "has", length(aliases), "aliases and",
        length(allegiances), "allegiances")
}
df %>% 
  pmap_chr(my_fun) %>% 
  tail()

pmap_chr(.l = df, .f = ~ my_fun(..1, ..2, ..3))
names(df)
```
